```{r}
library(tidyverse)
library(here)
library(fst)
library(webchem)
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)
library(conflicted)
library(fst)
library(ChemmineR)
library(VennDiagram)
library(wesanderson)
library(ggupset)
library(ggpubr)
library(tidyHeatmap)
library(ComplexHeatmap)
conflicts_prefer(select::dplyr)
conflicts_prefer(clusterProfiler::select)
conflicts_prefer(clusterProfiler::rename)
conflicts_prefer(clusterProfiler::filter)
conflicts_prefer(base::setdiff)
conflicts_prefer(tibble::view)
conflicts_prefer(dplyr::first)
conflicts_prefer(tidyHeatmap::heatmap)
```

# Prep kuster dose data

## Read in all 2-dose files and combine into one dataframe

Data was downloaded from the MASSIVE repository using the FTP client Filezilla at ftp://massive.ucsd.edu/v05/MSV000092248/ 
```{r}

all_files_path = here("data/new_kuster_data/per_dose_files")

all_files = list.files(all_files_path, full.names = TRUE)

all_data = data.frame()

for (i in 1:length(all_files)){
  
  this_file = all_files[i]

  this_drug_name = str_replace(basename(this_file), "_final\\.csv$", "")

  # Use tryCatch to handle errors
  this_file_data = tryCatch({
    # Attempt to read and process the file
    read_csv(this_file)  %>% 
      mutate(drug = this_drug_name)  %>% 
      filter(Picked.intensity != "Intensity") %>%
            # 2) …and drop those missing a DMSO variance
      filter(!is.na(Var.Normalized.LFQ.intensity.DMSO),
        # 2) …and drop those with 0 variance at 100nM or 1000nM
        Var.Normalized.LFQ.intensity.1000nM != 0,
        Var.Normalized.LFQ.intensity.100nM != 0) %>%
      # 4) but toss out the “full→zero→zero” patterns
      filter(
        !(Median.Normalized.LFQ.intensity.100nM == 0 &
          Median.Normalized.LFQ.intensity.1000nM == 0)
      ) %>%
      select(drug, gene = Gene.names, starts_with("Median.Normalized.LFQ.intensity")) %>%
      pivot_longer(starts_with("Median.Normalized.LFQ.intensity"), names_to = "dose", values_to = "intensity") %>%
      mutate(dose = gsub("Median.Normalized.LFQ.intensity.", "", dose))
  }, error = function(e) {
    # Handle errors by returning a custom message or action
    message("Data not appropriate type for file: ", this_file)
    return(NULL) # Return NULL or an appropriate value to avoid halting execution
  })

  if (!is.null(this_file_data)) {
    all_data = bind_rows(all_data, this_file_data)
  }
}
```

## Trametinib test
```{r}

trametinib_test = read_csv(here("data/new_kuster_data/per_dose_files/Trametinib_final.csv")) %>%
  mutate(drug = "Trametinib") %>%
      filter(Picked.intensity == "LFQ.intensity") %>%
      # 2) …and drop those missing a DMSO variance
      filter(!is.na(Var.Normalized.LFQ.intensity.DMSO),
        # 2) …and drop those with 0 variance at 100nM or 1000nM
    Var.Normalized.LFQ.intensity.1000nM != 0,
    Var.Normalized.LFQ.intensity.100nM != 0) %>%
      # 3) enforce non-increasing (keeps flats and true binders)
    #   filter(
    #    # Median.Normalized.LFQ.intensity.DMSO  >= Median.Normalized.LFQ.intensity.100nM,
    #     Median.Normalized.LFQ.intensity.100nM >= Median.Normalized.LFQ.intensity.1000nM
    #   ) %>%
      # 4) but toss out the “full→zero→zero” patterns
      filter(
        !(Median.Normalized.LFQ.intensity.100nM == 0 &
          Median.Normalized.LFQ.intensity.1000nM == 0)
      ) %>%
      # now tidy up
      mutate(drug) %>%
      select(drug, gene = Gene.names,
             starts_with("Median.Normalized.LFQ.intensity")) %>%
      pivot_longer(
        starts_with("Median.Normalized.LFQ.intensity"),
        names_to     = "dose",
        names_prefix = "Median.Normalized.LFQ.intensity.",
        values_to    = "intensity"
      )

```

## Read all full-dose files and combine into one dataframe

```{r}
full_dose_files = all_files %>% 
  as.data.frame()  %>% 
  rename(file = 1)  %>%
  filter(str_detect(file, "fulldose"))

full_dose_data = data.frame()

# 1) Define your dose columns
dose_cols <- c("0","0.3","1","3","10","30","100","300","1000")

for (i in 1:nrow(full_dose_files)){
  
  this_file = full_dose_files$file[i]

  this_drug_name = str_replace(basename(this_file), "_fulldose_final\\.csv$", "")

  this_file_data = read_csv(this_file)  %>%  
    # keep only the LFQ stream
    filter(usedData == "Normalized.LFQ.intensity.") %>%
    filter(down == TRUE) %>%
    # drop any row missing the 0 nM measurement
    filter(!is.na(`0`)) %>%
    ungroup() %>%
        select(drug = Drug,
            gene = Name,
            21:28) %>%
        pivot_longer(-c("drug", "gene"), names_to = "dose", values_to = "intensity") %>%
        mutate(dose = as.numeric(dose)*10^(-9))

    full_dose_data = bind_rows(full_dose_data, this_file_data)
  }
```

## Full dose test 
```{r}

# 1) Define your dose columns
dose_cols <- c("0","0.3","1","3","10","30","100","300","1000")

test <- read_csv(
    here("data/new_kuster_data/per_dose_files/SB-476429-A_fulldose_final.csv"),
    show_col_types = FALSE
  ) %>%
  # keep only the LFQ stream
  filter(usedData == "Normalized.LFQ.intensity.") %>%
  filter(down == TRUE) %>%
  # drop any row missing the 0 nM measurement
  filter(!is.na(`0`)) %>%
  ungroup() %>%
    select(drug = Drug,
           gene = Name,
           21:28) %>%
    pivot_longer(-c("drug", "gene"), names_to = "dose", values_to = "intensity") %>%
    mutate(dose = as.numeric(dose)*10^(-9))

```

## Standardize and prep data

```{r}

drugs = all_data %>% 
  select(drug) %>% 
  distinct()

doses = all_data %>% 
  select(dose) %>% 
  distinct()  %>% 
  arrange(dose)

data_processed = all_data  %>% 
    filter(dose != "DMSO")  %>% 
    mutate(dose = as.numeric(gsub("nM", "", dose))*10^-9)  %>%
    bind_rows(full_dose_data)  %>%
    filter(intensity < 3)  %>% 
    group_by(drug, gene, dose)  %>%
    summarize(relative_intensity = mean(intensity, na.rm = TRUE))  %>% 
    ungroup()  %>% 
    drop_na()

#distribution of doses
data_processed  %>% 
    ggplot(aes(x = dose)) +
    geom_histogram() +
    theme_minimal()

#distribution of relative intensities
data_processed  %>% 
    ggplot(aes(x = relative_intensity)) +
    geom_histogram() +
    theme_minimal()

processed_doses = data_processed %>%
  select(dose) %>% 
  distinct()  %>% 
  arrange(dose)

number_kinases = data_processed  %>% 
  select(gene) %>% 
  distinct()  %>% 
  nrow()

write.fst(data_processed, here("results/processed_kuster_dose_data_LFQ_only.fst"))
```

# Combine with LINCS, klaeger, and okl data

```{r}
LINCS_data = read_csv(here("data/LINCS_data/full_LINCS_UNC_1uM_data_imputed.csv"))

LINCS_data_processed = LINCS_data  %>% 
    pivot_longer(-c("Small.Molecule.Name", "Dose"), names_to = "gene", values_to = "percent_control")  %>% 
    select(drug = Small.Molecule.Name,
            dose = Dose,
            gene,
            relative_intensity = percent_control)  %>%
    mutate(dose = dose*10^-6, 
           relative_intensity = relative_intensity/100) 

klaeger_data = read_rds(here("data/klaeger_data/klaeger_full_tidy.rds"))  %>% 
    select(drug, gene = gene_name, dose = concentration_M, relative_intensity)

okl_data = read_rds(here("results/okl_profiling_data.rds.gz"))  %>% 
    select(drug, gene = discoverx_gene_symbol, entrez_gene_symbol, dose = concentration_nM, relative_intensity = percent_control)
  
okl_data_processed = okl_data  %>% 
    mutate(dose = dose*10^-9, 
           relative_intensity = relative_intensity/100)

```

# Prep combined data

## Combine drug info
```{r}

LINCS_cids = read_csv(here("data/matching/all_LINCS_cids.csv"))  %>% 
  mutate(cid = as.character(cid))  %>% 
  select(Compound = LINCS_name, cid)

klaeger_compound_info = read_csv(here("data/klaeger_data/klaeger_compound_info.csv"))  %>% 
    select(Compound = Drug, SMILES = `SMILES code`, cid = PubChemCID)  %>% 
    mutate(cid = as.character(cid))

kuster_compound_info_tool = read_csv(here("data/new_kuster_data/compound_info.csv"))
kuster_compound_info_clinical = read_csv(here("data/new_kuster_data/compound_info_clinical.csv"))
kuster_compound_info = bind_rows(kuster_compound_info_tool, kuster_compound_info_clinical  %>% rename(Compound = Drug))  %>% 
    select(Compound, SMILES)
kuster_compound_cid = get_cid(unique(kuster_compound_info$SMILES), from = "smiles", domain = "compound", match = "all", verbose = T)
kuster_compound_info = kuster_compound_info  %>% 
    left_join(kuster_compound_cid, by = c("SMILES" = "query"))  %>% 
    select(Compound, SMILES, cid)

okl_compound_info = read_fst(here("data/okl_profiling_data/lsp_compound_dictionary.fst"))  %>% 
    select(Compound = pref_name, inchi)  %>% 
    distinct()  %>%
    filter(Compound %in% okl_data$drug)  %>%
    drop_na()
okl_cid = pubchemInchi2cid(okl_compound_info$inchi  %>% unique())
okl_compound_info = okl_compound_info  %>% 
    select(Compound, inchi)  %>%
    unique()  %>% 
    bind_cols(cid = okl_cid)  %>% 
    mutate(cid = as.character(cid))
okl_0_cid = okl_compound_info  %>% 
    filter(is.na(cid) | cid == 0)  %>% 
    select(Compound)  %>% 
    distinct()
okl_0_cid = get_cid(okl_0_cid$Compound, from = "name", domain = "compound", match = "all", verbose = T)
okl_compound_info = okl_compound_info  %>% 
    left_join(okl_0_cid  %>% rename(cid_name = cid), by = c("Compound" = "query"))  %>% 
    mutate(cid = if_else(is.na(cid) | cid == 0, cid_name, cid))  %>%
    select(Compound, inchi, cid)

all_compound_info = bind_rows(kuster_compound_info, klaeger_compound_info, okl_compound_info, LINCS_cids)  %>% 
  select(drug = Compound, cid, SMILES, inchi)  %>% 
  distinct()
```

## Combine data
```{r}
kuster_dose_data = read_fst(here("results/processed_kuster_dose_data_LFQ_only.fst"))

combined_data = bind_rows(kuster_dose_data  %>% mutate(dataset = "kuster"),
                          LINCS_data_processed  %>% mutate(dataset = "LINCS"),
                          klaeger_data  %>% mutate(dataset = "klaeger"),
                          okl_data_processed  %>% mutate(dataset = "okl"))

combined_compound_info = combined_data  %>% 
    select(drug, dataset)  %>% 
    distinct()  %>% 
    left_join(all_compound_info, by = "drug", relationship = "many-to-many")  %>% 
    distinct()  %>% 
    mutate(cid = if_else(str_detect(drug, "Lestaurtinib"), "126565", cid))  %>% 
    filter(!is.na(cid) | !is.na(SMILES) | !is.na(inchi))  %>% 
    select(-inchi)

combined_dataset_compound_info = combined_data  %>%
    left_join(combined_compound_info, by = "drug", relationship = "many-to-many")  %>%
    distinct()  %>%
    mutate(cid = if_else(is.na(cid) | cid == 0, drug, cid))

#all_kinase_info is from prepping the kd_data and manually looking up the kinase names (check src/data_organization/prep_all_kd_data.Rmd)
kinase_info = read_csv(here("results/all_kinase_info.csv"))  %>% 
    mutate(ENTREZID = as.character(ENTREZID))

combined_kinase_info = combined_dataset_compound_info  %>%
    select(gene)  %>% 
    distinct() %>%  
    left_join(kinase_info, by = c("gene" = "original_kinase_name")) %>%
    drop_na()

# na_target = combined_kinase_info  %>% 
#     filter(is.na(target))  %>% 
#     select(gene)  %>% 
#     distinct()

# na_target_entrez = bitr(na_target$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = org.Hs.eg.db)

# combined_kinase_info = combined_kinase_info  %>% 
#     drop_na()  %>%
#     bind_rows(na_target_entrez  %>%
#                   rename(gene = SYMBOL)  %>%  
#                   mutate(target = gene))

combined_dataset_compound_kinase_info = combined_dataset_compound_info  %>%
    left_join(combined_kinase_info, by = "gene")

kinase_frequency = combined_dataset_compound_kinase_info  %>% 
    count(target)

combined_dataset_final = combined_dataset_compound_kinase_info  %>% 
    filter(target %in% kinase_frequency$target)  %>% 
    filter(!is.na(target))  %>%
    distinct()  %>%  
    select(compound = drug,
            compound_dose_M = dose,
            compound_pubchem_cid = cid,
            compound_smiles = SMILES,
            dataset = dataset.x,
            original_target_name = gene,
            target,
            target_entrez_id = ENTREZID,
            relative_intensity)

write.fst(combined_dataset_final, here("results/combined_dose_data_LFQ_only.fst"))

```

## EDA
```{r}
combined_dataset = read_fst(here("results/combined_dose_data_LFQ_only.fst"))

#number kinases
combined_dataset  %>% 
    select(target) %>% 
    distinct()  %>% 
    nrow()

#number compounds
combined_dataset  %>% 
    select(compound_pubchem_cid)  %>% 
    distinct()  %>% 
    nrow()

#doses of compounds
doses = combined_dataset  %>%
    mutate(compound_dose_M  = as.numeric(as.character(compound_dose_M)))  %>% 
	mutate(compound_dose_uM = compound_dose_M * 10^6)  %>%
	select(-compound_dose_M)  %>%
    filter(compound_dose_uM != 0)  %>% 
    count(compound_dose_uM)

#distribution of mean relative intensities
combined_dataset  %>% 
    ggplot(aes(x = relative_intensity)) +
    geom_histogram() +
    theme_minimal()

```

## Make Venn Diagram for compounds in each dataset
```{r}
combined_dataset = read_fst(here("results/combined_dose_data_LFQ_only.fst"))

dataset_for_plot = combined_dataset  %>% 
    select(compound_pubchem_cid, dataset)  %>% 
    distinct()  %>% 
    filter(compound_pubchem_cid != 0)  %>% 
    drop_na()

# Split the dataset by the 'dataset' column to get unique compounds from each original dataset
datasets_split <- split(dataset_for_plot, dataset_for_plot$dataset)

# Get unique compounds for each dataset
unique_compounds <- lapply(datasets_split, function(df) unique(df$compound_pubchem_cid))

# Prepare a named list for VennDiagram 
list_for_venn <- setNames(unique_compounds, names(unique_compounds))

#use wesanderson palette
myCol <- wes_palette("Zissou1", 4)

# Generate and plot the Venn diagram (for 4 datasets, a quad Venn diagram)
venn.plot <- venn.diagram(
  x = list_for_venn,
  filename = here("figures/venn_diagram_dose_data_LFQ_only.png"),
  category.names = names(list_for_venn),
  output = TRUE,
    # Output features
    imagetype="png" ,
    height = 1480 , 
    width = 1480 , 
    resolution = 300,
  #plot features
  lwd = 2,
  lty = "blank",
  fill = myCol,
  cex = .6,
  fontface = "bold",
  fontfamily = "sans"
)
```

# Make UpSet plot for compounds in each dataset
```{r}
upset_data = combined_dataset  %>% 
    select(compound_pubchem_cid, dataset)  %>% 
    distinct()  %>% 
    filter(compound_pubchem_cid != 0)  %>% 
    group_by(compound_pubchem_cid)  %>%
    summarise(dataset = list(dataset))  %>%
    drop_na()

total_drugs = combined_dataset  %>% 
    select(compound_pubchem_cid)  %>% 
    distinct()  %>% 
    nrow()

# Create the UpSet plot
p <- ggplot(upset_data) +
  geom_bar(aes(x = dataset), fill = "steelblue") +
  geom_text(stat='count', aes(x = dataset, label=after_stat(count)), vjust=-1) +
  expand_limits(y = c(0, 780)) +
  scale_x_upset(n_intersections = 10) +
  labs(x = "Compound PubChem CID", y = "Number of Datasets", title = "Unique Compounds in Each Kinome State Dataset", subtitle = paste0("Total Unique Compounds: ", total_drugs)) +
  theme_pubr()

ggsave(here("figures/upset_plot_combined_dose_data_compounds_LFQ_only.png"), p, width = 8, height = 5, dpi = 300)

```

# Make UpSet plot for targets in each dataset
```{r}
upset_data = combined_dataset  %>% 
    select(target, dataset)  %>% 
    distinct()  %>% 
    filter(!is.na(target))  %>% 
    group_by(target)  %>%
    summarise(dataset = list(dataset))  %>%
    drop_na()

total_targets = combined_dataset  %>% 
    select(target)  %>% 
    distinct()  %>% 
    nrow()

# Create the UpSet plot
p <- ggplot(upset_data) +
  geom_bar(aes(x = dataset), fill = "#950505") +
    geom_text(stat='count', aes(x = dataset, label=after_stat(count)), vjust=-1) +
    labs(x = "Target", y = "Number of Datasets", title = "Unique Targets in Each Kinome State Dataset", subtitle = paste0("Total Unique Targets: ", total_targets)) +
    expand_limits(y = c(0, 220)) +
  scale_x_upset(n_intersections = 10) +
  theme_pubr()
ggsave(here("figures/upset_plot_combined_dose_data_targets_LFQ_only.png"), p, width = 8, height = 5, dpi = 300)
```

### Make Heatmap of trametinib at all doses for each dataset
```{r}
combined_dataset = read_fst(here("results/combined_dose_data_LFQ_only.fst"))

heatmap_data = combined_dataset  %>% 
    filter(compound_pubchem_cid == "11707110")  %>% 
    select(compound_dose_M, dataset, target, relative_intensity)  %>%
    mutate(compound_dose_M = as.numeric(as.character(compound_dose_M)))  %>% 
    mutate(compound_dose_uM = compound_dose_M * 10^6)  %>%
    mutate(dataset_dose = paste0(dataset, "_", compound_dose_uM))  %>%
    select(-compound_dose_M)  %>%
    filter(compound_dose_uM != 0)  %>%
    drop_na()

ht = heatmap_data %>%
    as_tibble() %>%
  heatmap(
    .row = dataset_dose,
    .column = target,
    .value = relative_intensity,
    scale = "none",
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    #cluster_columns = FALSE,
    palette_value = circlize::colorRamp2(
      seq(0, 1, length.out = 11), 
      rev(RColorBrewer::brewer.pal(11, "RdBu"))
    ),
    row_names_gp = gpar(fontsize = 10),
    column_names_gp = gpar(fontsize = 4),
    column_title = "Trametinib Specificity at All Doses",
    show_column_names = TRUE,
    heatmap_legend_param = list(title = "Relative Intensity", title_position = "topcenter", legend_direction = "horizontal")
  ) 


png("figures/heatmap_trametinib_all_doses_LFQ_only.png", width = 5500, height = 1000, res = 300)
# Draw the heatmap
draw(
  ht %>% as_ComplexHeatmap(),
  heatmap_legend_side = "bottom",
  legend_grouping = "original",
  padding = unit(c(10, 2, 2, 2), "mm")
)
dev.off()

ht

```

### Make Heatmap of erlotinib at all doses for each dataset
```{r}
combined_dataset = read_fst(here("results/combined_dose_data_LFQ_only.fst"))

heatmap_data = combined_dataset  %>% 
    filter(compound_pubchem_cid == "176870")  %>% 
    select(compound_dose_M, dataset, target, relative_intensity)  %>%
    mutate(compound_dose_M = as.numeric(as.character(compound_dose_M)))  %>% 
    mutate(compound_dose_uM = compound_dose_M * 10^6)  %>%
    mutate(dataset_dose = paste0(dataset, "_", compound_dose_uM))  %>%
    select(-compound_dose_M)  %>%
    filter(compound_dose_uM != 0)  %>%
    drop_na()

ht = heatmap_data %>%
    as_tibble() %>%
    filter(target == "EGFR") %>%
  heatmap(
    .row = dataset_dose,
    .column = target,
    .value = relative_intensity,
    scale = "none",
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    #cluster_columns = FALSE,
    palette_value = circlize::colorRamp2(
      seq(0, 1, length.out = 11), 
      rev(RColorBrewer::brewer.pal(11, "RdBu"))
    ),
    row_names_gp = gpar(fontsize = 10),
    column_names_gp = gpar(fontsize = 4),
    column_title = "Erlotinib Specificity at All Doses",
    show_column_names = TRUE,
    heatmap_legend_param = list(title = "Relative Intensity", title_position = "topcenter", legend_direction = "horizontal")
  ) 


png("figures/heatmap_erlotinib_all_doses_LFQ_only.png", width = 5500, height = 1000, res = 300)
# Draw the heatmap
draw(
  ht %>% as_ComplexHeatmap(),
  heatmap_legend_side = "bottom",
  legend_grouping = "original",
  padding = unit(c(10, 2, 2, 2), "mm")
)
dev.off()

ht

```

### Make Heatmap of all drugs at 1 uM (all kinases, missing imputed to 0)
```{r}
combined_dataset_summarized = read_fst(here("results/combined_dose_data_summarized_LFQ_only.fst"))

heatmap_all_1uM = combined_dataset_summarized %>% 
    mutate(
        compound_dose_uM = as.numeric(as.character(compound_dose_uM)),
        target_entrez_id = as.character(target_entrez_id)
    ) %>% 
    filter(!is.na(compound_dose_uM), compound_dose_uM == 1) %>% 
    drop_na(compound_pubchem_cid, compound_pubchem_name, target_entrez_id)

# summarise to one value per compound/target
heatmap_all_1uM = heatmap_all_1uM %>% 
    group_by(compound_pubchem_name, compound_pubchem_cid, target_entrez_id) %>% 
    summarise(percent_inhibition = mean(percent_inhibition, na.rm = TRUE), .groups = "drop")

# build full compound x target grid and impute missing percent_inhibition to 0 (faster than complete)
compounds_1uM = heatmap_all_1uM %>% distinct(compound_pubchem_name, compound_pubchem_cid)
targets_1uM = heatmap_all_1uM %>% distinct(target_entrez_id)

heatmap_all_1uM_completed = tidyr::crossing(compounds_1uM, targets_1uM) %>%
    left_join(heatmap_all_1uM, by = c("compound_pubchem_name", "compound_pubchem_cid", "target_entrez_id")) %>%
    mutate(
        percent_inhibition = coalesce(percent_inhibition, 0),
        row_label = paste0(compound_pubchem_name, " (", compound_pubchem_cid, ")")
    )

# map entrez -> symbol for nicer columns
entrez_to_symbol = heatmap_all_1uM_completed %>% 
    distinct(target_entrez_id) %>% 
    pull(target_entrez_id) %>% 
    bitr(fromType = "ENTREZID", toType = "SYMBOL", OrgDb = org.Hs.eg.db) %>%
    rename(target_entrez_id = ENTREZID, target_symbol = SYMBOL)

heatmap_all_1uM_labeled = heatmap_all_1uM_completed %>% 
    left_join(entrez_to_symbol, by = "target_entrez_id") %>% 
    mutate(target_symbol = if_else(is.na(target_symbol), target_entrez_id, target_symbol))

ht_all = heatmap_all_1uM_labeled %>%
    as_tibble() %>%
    heatmap(
    .row = row_label,
    .column = target_symbol,
    .value = percent_inhibition,
    scale = "none",
    cluster_rows = TRUE,
    cluster_columns = TRUE,
    palette_value = circlize::colorRamp2(
        seq(0, 1, length.out = 9), 
        RColorBrewer::brewer.pal(9, "YlGnBu")
    ),
    row_names_gp = gpar(fontsize = 10),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Percent Inhibition at 1 uM (All Kinases, Missing=0)",
    show_column_names = TRUE,
    show_row_names = FALSE,
    heatmap_legend_param = list(title = "Percent Inhibition", title_position = "topcenter", legend_direction = "horizontal")
    )

png("figures/heatmap_all_drugs_1uM_all_targets_LFQ_only.png", width = 7500, height = 7500, res = 300)
draw(
    ht_all %>% as_ComplexHeatmap(),
    heatmap_legend_side = "bottom",
    legend_grouping = "original",
    padding = unit(c(10, 2, 2, 2), "mm")
)
dev.off()

ht_all

```


# Writing out summarized dataset

Excluding the OKL data for now, since it is not yet published by the original authors and we don't have permission to share it.

```{r}
combined_dataset = read_fst(here("results/combined_dose_data_LFQ_only.fst"))

all_compounds = combined_dataset  %>% 
    select(compound, compound_pubchem_cid)  %>% 
    distinct()  %>% 
    mutate(compound_pubchem_cid = as.numeric(compound_pubchem_cid))  %>% 
    drop_na()
  
unique_cids <- unique(all_compounds$compound_pubchem_cid)

# Updated function with error handling
get_readable_synonym <- function(cid) {

synonyms = pc_synonyms(cid, from = "cid", verbose = TRUE)

first_synonym = synonyms %>% 
    pluck(1)  %>% 
    pluck(1)

return(as.character(first_synonym))

}

# Use map_chr to apply the function and ensure character output
readable_synonyms <- map_chr(unique_cids, get_readable_synonym)

# Create the data frame
readable_synonyms_df <- data.frame(cid = unique_cids, pubchem_name = readable_synonyms, stringsAsFactors = FALSE)

write_csv(readable_synonyms_df, here("results/combined_dose_compound_synonyms_LFQ_only.csv"))
readable_synonyms_df = read_csv(here("results/combined_dose_compound_synonyms_LFQ_only.csv"))  %>% 
    filter(!is.na(cid))  %>%
    filter(cid != 0)  %>% 
    mutate(cid = as.character(cid))

compound_cid_info_smiles = pc_prop(all_compounds$compound_pubchem_cid  %>% unique(), properties = c("CanonicalSMILES"))  %>% 
    mutate(CID = as.character(CID))

combined_dataset_compound_info = combined_dataset  %>%
    mutate(compound_pubchem_cid = as.character(compound_pubchem_cid))  %>%
    left_join(readable_synonyms_df, by = c("compound_pubchem_cid" = "cid"))  %>%
    rename(compound_pubchem_name = pubchem_name)  %>%
    left_join(compound_cid_info_smiles, by = c("compound_pubchem_cid" = "CID"))  %>%
    mutate(compound_pubchem_name = case_when(
        is.na(compound_pubchem_name) ~ compound,
        TRUE ~ compound_pubchem_name
    ),
    compound_smiles = case_when(
        is.na(compound_smiles) ~ ConnectivitySMILES,
        TRUE ~ compound_smiles
    ))  %>% 
    select(-ConnectivitySMILES)

combined_dataset_summarized_no_okl = combined_dataset_compound_info  %>% 
    filter(dataset != "okl")  %>%
    mutate(compound_dose_M  = as.numeric(as.character(compound_dose_M)))  %>% 
	mutate(compound_dose_uM = compound_dose_M * 10^6)  %>%
	select(-compound_dose_M)  %>%
    filter(compound_dose_uM != 0)  %>%
    #impute >1 values as 1
    mutate(relative_intensity = if_else(relative_intensity > 1, 1, relative_intensity))  %>%
    group_by(compound_pubchem_name, compound_pubchem_cid, compound_smiles, compound_dose_uM, target, target_entrez_id)  %>%
    summarise(mean_relative_intensity = mean(relative_intensity, na.rm = T))  %>% 
    ungroup()  %>%
    mutate(percent_inhibition = 1 - mean_relative_intensity)  %>%
    select(-mean_relative_intensity)  %>%
    write_fst(here("results/combined_dose_data_summarized_no_okl_LFQ_only.fst"))

#write out csv for app
data_for_app = combined_dataset_summarized_no_okl  %>%
    group_by(compound_pubchem_name, compound_dose_uM, target, target_entrez_id)  %>%
    summarise(percent_inhibition = mean(percent_inhibition, na.rm = T))  %>%
    ungroup()  %>%
    write_csv(here("results/combined_dose_data_for_app_LFQ_only.csv"))

# distribution of percent inhibition
data_for_app  %>% 
    ggplot(aes(x = percent_inhibition)) +
    geom_histogram() +
    theme_minimal()

combined_dataset_summarized = combined_dataset_compound_info  %>% 
    mutate(compound_dose_M  = as.numeric(as.character(compound_dose_M)))  %>% 
	mutate(compound_dose_uM = compound_dose_M * 10^6)  %>%
	select(-compound_dose_M)  %>%
    filter(compound_dose_uM != 0)  %>%
    #impute >1 values as 1
    mutate(relative_intensity = if_else(relative_intensity > 1, 1, relative_intensity))  %>%
    group_by(compound_pubchem_name, compound_pubchem_cid, compound_smiles, compound_dose_uM, target, target_entrez_id)  %>%
    summarise(mean_relative_intensity = mean(relative_intensity, na.rm = T))  %>% 
    ungroup()  %>%
    mutate(percent_inhibition = 1 - mean_relative_intensity)  %>%
    select(-mean_relative_intensity)  %>%
    write_fst(here("results/combined_dose_data_summarized_LFQ_only.fst"))

paclitaxel = combined_dataset_summarized  %>% 
    filter(compound_pubchem_cid =="36314")
```

### Make Heatmap of trametinib (summarized data)
```{r}
combined_dataset_summarized = read_fst(here("results/combined_dose_data_summarized_LFQ_only.fst"))

heatmap_data_summarized = combined_dataset_summarized  %>% 
        filter(compound_pubchem_cid == "11707110")  %>% 
        select(compound_dose_uM, target, percent_inhibition)  %>%
        mutate(compound_dose_uM = as.numeric(as.character(compound_dose_uM)))  %>% 
        filter(compound_dose_uM != 0)  %>% 
        drop_na()

ht_summarized = heatmap_data_summarized %>%
        mutate(dose_label = paste0("dose_", compound_dose_uM, "uM")) %>%
        as_tibble() %>%
    heatmap(
        .row = dose_label,
        .column = target,
        .value = percent_inhibition,
        scale = "none",
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        palette_value = circlize::colorRamp2(
            seq(0, 1, length.out = 9), 
            RColorBrewer::brewer.pal(9, "YlGnBu")
        ),
        row_names_gp = gpar(fontsize = 10),
        column_names_gp = gpar(fontsize = 4),
        column_title = "Trametinib Percent Inhibition (Summarized)",
        show_column_names = TRUE,
        heatmap_legend_param = list(title = "Percent Inhibition", title_position = "topcenter", legend_direction = "horizontal")
    ) 

png("figures/heatmap_trametinib_summarized_LFQ_only.png", width = 5500, height = 1000, res = 300)
draw(
    ht_summarized %>% as_ComplexHeatmap(),
    heatmap_legend_side = "bottom",
    legend_grouping = "original",
    padding = unit(c(10, 2, 2, 2), "mm")
)
dev.off()

ht_summarized

```

## Add non-kinase inhibior target data for novartis

```{r}
novartis_targets = read_csv(here("data/novartis_drug_target_kds.csv"))  %>% 
    mutate(cid = if_else(novartis_name == "dacarbazine", 135398738, cid))

novartis_entrez = bitr(novartis_targets$uniprot_id, fromType = "UNIPROT", toType = "ENTREZID", OrgDb = org.Hs.eg.db)
novartis_symbol = bitr(novartis_targets$uniprot_id, fromType = "UNIPROT", toType = "SYMBOL", OrgDb = org.Hs.eg.db)
novartis_cid_info = readable_synonyms <- map_chr(novartis_targets$cid  %>% unique(), get_readable_synonym)
novartis_cid_info = data.frame(cid = novartis_targets$cid  %>% unique(), pubchem_name = readable_synonyms, stringsAsFactors = FALSE)  %>% 
    filter(!is.na(cid))  %>%
    filter(cid != 0)  %>% 
    mutate(cid = as.character(cid))

novartis_targets_entrez = novartis_targets  %>%
    mutate(cid = as.character(cid))  %>%
    left_join(novartis_entrez, by = c("uniprot_id" = "UNIPROT"))  %>% 
    left_join(novartis_symbol, by = c("uniprot_id" = "UNIPROT"))  %>%
    left_join(novartis_cid_info, by = c("cid"))  %>%
    mutate(pubchem_name = if_else(is.na(pubchem_name), novartis_name, pubchem_name))  %>%
    mutate(cid = if_else(is.na(cid), novartis_name, cid))  %>%
    select(target = SYMBOL, target_entrez_id = ENTREZID, compound_pubchem_cid = cid, compound_pubchem_name = pubchem_name, compound_smiles = CanonicalSMILES, mean_kd_apparent = kd)  %>% 
    #convert kd to "fraction bound" at 1 uM (1000 nM because Kds are in nM)
    mutate(percent_inhibition = (1000 / (1000 + mean_kd_apparent)), compound_dose_uM = 1)  %>% 
    select(-mean_kd_apparent)

# combine with the rest of the data
combined_dataset_summarized = read_fst(here("results/combined_dose_data_summarized.fst"))  %>% 
    mutate(target_entrez_id = as.character(target_entrez_id))  %>%
    bind_rows(novartis_targets_entrez)  %>% 
    write_fst(here("results/combined_novartis_dose_data_summarized.fst"))

combined_dataset_summarized_no_okl = read_fst(here("results/combined_dose_data_summarized_no_okl.fst"))  %>% 
    mutate(target_entrez_id = as.character(target_entrez_id))  %>%
    bind_rows(novartis_targets_entrez)  %>% 
    write_fst(here("results/combined_novartis_dose_data_summarized_no_okl.fst"))  %>% 
    write_csv(here("results/combined_novartis_dose_data_summarized_no_okl.csv"))

test_for_duplicates = combined_dataset_summarized_no_okl  %>% 
    count(compound_pubchem_cid, target, compound_dose_uM)

palbociclib = combined_dataset_summarized  %>% 
    filter(compound_pubchem_cid == "5330286")
```

# Getting kinase diversity lists for CORAL

```{r}
combined_dataset = read_fst(here("results/combined_dose_data_summarized_no_okl.fst"))

KCGS_kinases = combined_dataset  %>%
    filter(dataset == "LINCS")  %>%
    filter(relative_intensity < 0.01)  %>%
    select(target_entrez_id)  %>% 
    distinct()  %>% 
    mutate(group = "Hit_>_80%")

All_kinases = combined_dataset  %>%
    filter(relative_intensity < 0.2)  %>%
    select(target_entrez_id)  %>% 
    distinct()  %>% 
    mutate(group = "Hit_>_80%")

library(clipr)

KCGS_kinases  %>% 
    write_clip()

All_kinases  %>%
    write_clip()

#number kinases
combined_dataset  %>% 
    select(target_entrez_id)  %>% 
    distinct()  %>% 
    nrow()

#number compounds
combined_dataset  %>% 
    select(compound_pubchem_cid)  %>% 
    distinct()  %>% 
    nrow()

#doses of compounds
combined_dataset  %>% 
    select(compound_pubchem_cid, compound_dose_uM)  %>% 
    distinct()  %>% 
    arrange(compound_dose_uM)

```